---
title: "p8105_hw2_hx2328"
author: "Hao Xu"
date: "10/8/2021"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

# Problem 1

Read and clean the Mr. Trash Wheel data

```{r}
trash_wheel_clean = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls))
```

Read and clean 2018 and 2019 precipitation data.

```{r}
precip_2018 = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018)

precip_2019 = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 4, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2019)
```

Combine precipitation dataset and convert month to a character variable

```{r}
precip_data = bind_rows(precip_2018, precip_2019) %>% 
  mutate(month = month.name[month])
```

The key variables in the Mr. Trash Wheel are the weight_towns, and the month and year. We could find the relationship between the weight of trash and precipitation amount.

If we calculate the total precipitation in 2018 and median number of sports balls in 2019:

```{r}
total_2018 = filter(precip_data, year == 2018)
sports_balls_2019 = filter(trash_wheel_clean, year == 2019)
```

The total precipitation in 2018 would be `r sum(total_2018$total)`.

The median number of sports balls in 2019 would be `r median(sports_balls_2019$sports_balls)`


# Problem 2

Read and clean pols-month.csv data.

```{r}
pols_month = read_csv("fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = prez_dem,
         president = recode(president, `1` = "dem", `0` = "gop"),
         president = factor(president),
         year = as.numeric(year)) %>% 
  select(-day, -prez_gop, -prez_dem)
```

Read and clean snp.csv data.

```{r}
snp = read_csv("fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(month = month.name[as.numeric(month)],
         year = as.numeric(year))

#There is something wrong with the date format when I read the csv file. 
#The year variable is 2 digit instead of 4, so did some modifications by loop.
for (i in 1:787){
  if (snp$year[i] <= 15){
    snp$year[i] = snp$year[i] + 2000
  }
  else{
    snp$year[i] = snp$year[i] + 1900
  }
}

snp = select(snp, -day) %>% 
  arrange(year, month) %>% 
  relocate(year, month)
```

Read and clean the unemployment.csv data.

```{r}
unemployment = read_csv("fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  mutate(year = as.numeric(year)) %>% 
  pivot_longer(jan:dec, names_to = "month", values_to = "percentage_unemployment") %>% 
  mutate(month = recode(month,
                        "jan" = "January",
                        "feb" = "February",
                        "mar" = "March",
                        "apr" = "April",
                        "may" = "May",
                        "jun" = "June",
                        "jul" = "July",
                        "aug" = "August",
                        "sep" = "September",
                        "oct" = "October",
                        "nov" = "November",
                        "dec" = "December"
                        ))
```

Merge data:

```{r}
merged_data = left_join(pols_month, snp) %>% 
  left_join(unemployment)
```

The snp contains year, month, and the closing values of the S&P stock index;

The pols mainly contains the year, month, and the political status of the president.

The merged dataset has dimension `r nrow(merged_data)` * `r ncol(merged_data)`. The range of years is `r range(merged_data$year)`. The names of the key variables include: `r colnames(merged_data)`.


# Problem 3

Read data and clean the variable name

```{r}
popular_names = read_csv("Popular_Baby_Names.csv") %>% 
  janitor::clean_names()
```

Check if there are changes of the names of a categorical predictor and the case structure of string variables.

```{r}
table(popular_names$year_of_birth) #good
table(popular_names$gender)#good
table(popular_names$ethnicity)#detected some changes
```

Address the problem and omit duplicate rows:

```{r}
popular_names_clean = mutate(popular_names, ethnicity = recode(ethnicity, 
                                                               "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                                                               "BLACK NON HISP" = "BLACK NON HISPANIC",
                                                               "WHITE NON HISP" = "WHITE NON HISPANIC")) %>% 
  unique()
```

Produce the "Olivia" as female baby name rank table.

```{r}
olivia_female = filter(popular_names_clean, gender == "FEMALE", childs_first_name == "Olivia") %>% 
  select(ethnicity, year_of_birth, rank) %>% 
  arrange(ethnicity, year_of_birth)
olivia_table = matrix(olivia_female$rank, ncol = 4)
colnames(olivia_table) = c(2013, 2014, 2015, 2016)
rownames(olivia_table) = c("ASIAN AND PACIFIC ISLANDER", "BLACK NON HISPANIC", "HISPANIC", "WHITE NON HISPANIC")
as.table(olivia_table)
```

Produce another table showing the most famous male name over the time.

```{r}
male_name = filter(popular_names_clean, gender == "MALE", rank == 1) %>% 
  arrange(year_of_birth)
male_table = matrix(male_name$childs_first_name, ncol = 6)
colnames(male_table) = c(2011:2016)
rownames(male_table) = c("ASIAN AND PACIFIC ISLANDER", "BLACK NON HISPANIC", "HISPANIC", "WHITE NON HISPANIC")
as.table(male_table)
```

Finally, got the scatter plot for male, white non-hispanic children born in 2016, show the number of children with a name against the rank in popularity of that name.

```{r}
male_white = filter(popular_names_clean, gender == "MALE", ethnicity == "WHITE NON HISPANIC",
                    year_of_birth == 2016)
ggplot(male_white, aes(x = rank, y = count)) + geom_point() + xlab("The rank in popularity of that name") + 
  ylab("The number of children with a name")
ggsave("scatter_plot.pdf", height = 4, width = 6)
```

